---
title: "SBIR Awards Summary"
author: "Logan Kocka"
date: "6/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(plotly)
library(data.table)
library(tibble)
```

User inputs
```{r, message=FALSE}
fileSource <- "NSF" #used in title of plots
numPrint <- 10 #prints this many of the highest below plots
cutoffYear <- 2010
mostRecent <- "Jun 2020" #most recent date of pulled data 
titleSnippet <- paste0(fileSource, ", ", cutoffYear, " to ", mostRecent)
```

Load in the data and sort by Organization name.
```{r, message=FALSE}
data <- read.csv('SBIR 2020-06-08 armi_df_whole.csv', stringsAsFactors = F) #choose file
data <- data[order(data$Organization),] #order according to Organization (A-Z)
```

Clean the data: remove commas, "Inc.", "Inc", "Llc", and trailing white spaces.
```{r, message=FALSE}
totalRows <- nrow(data)
cleaned <- data
cleaned$Organization <- gsub(",", "", cleaned$Organization) #removes commas
cleaned$Organization <- gsub("Incorporated", "", cleaned$Organization) #removes Incorp
cleaned$Organization <- gsub("Inc.", "", cleaned$Organization) #removes Inc.
cleaned$Organization <- gsub("Inc", "", cleaned$Organization) #removes Inc
cleaned$Organization <- gsub("Llc", "", cleaned$Organization) #removes Llc
cleaned$Organization <- trimws(cleaned$Organization) #removes trailing spaces
cleaned$Organization <- gsub("\\s+", " ", cleaned$Organization) #merges multiple white spaces into one
cleaned_cutoff <- cleaned %>% filter(cleaned$FY >= cutoffYear)

print(paste0(totalRows, " total unique awards"))
print(paste0(nrow(cleaned %>% filter(cleaned$FY == 0)), " did not capture a year"))
print(paste0(nrow(cleaned_cutoff), " were awarded after the cutoff year"))
print(paste0(nrow(cleaned_cutoff %>% filter(cleaned_cutoff$Funding == 0)), " showed $0 in funding"))
```

Stacked histogram - Total number of awards per organization
```{r, message=FALSE, warning=F}
filtered <- cleaned_cutoff %>% select(Organization, Tissue) #select rows you want to keep
counts <- as.data.frame(table(filtered), stringsAsFactors = F) 
#consolidate by tissue type and count frequency
counts <- counts %>% filter(counts$Freq != 0) #removes frequencies of zero

#order bars in plot
#remove tissue type column  #sum freq's per org
countsOrder <- counts %>% select(!Tissue) %>% group_by(Organization) %>% summarize(Freq = sum(Freq))
#order by Freq to establish customer factors
countsOrder <- countsOrder[order(countsOrder$Freq, decreasing=F),]
counts$Organization <- factor(counts$Organization, level=countsOrder$Organization)

#create stacked histogram plot showing breakdown of awards per organization and their tissue types
hist <- plot_ly(y = ~counts$Organization, x = ~counts$Freq, type = 'bar', orientation = 'h', 
                name = ~counts$Tissue, color = ~counts$Tissue) %>%
  layout(margin = list(l = 300), title=paste0("Tissue Type / Organization Breakdown (", titleSnippet, ")"), 
         yaxis = list(title = 'Number of Awards'), xaxis = list(title = 'Organization Name'), barmode = 'stack')
hist
```

Top 10 organizations by most total awards
```{r}
for(i in 0:numPrint){
  print(as.character(countsOrder$Organization[length(countsOrder$Organization)-i]))
}
```

Pie plot - Total number of awards per tissue type
```{r}
#create pie plot showing breakdown of total number of awards per tissue type
pieData <- as.data.frame(table(filtered[,-1]), stringsAsFactors = F)
pie <- plot_ly(pieData, labels = ~Var1, values = ~Freq, type = 'pie') %>% 
  layout(title = paste0("Tissue Distribution (", titleSnippet, ")"),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie
```
Top 5 tissue types by number of awards
```{r}
pieData <- pieData[order(pieData$Freq, decreasing = T),] #order DF by frequency
for(i in 1:5){ #print top 5 tissue types 
  print(pieData$Var1[i])
}
```

Bar plot - Amount of funding per organization
```{r, message=FALSE, warning=F}
#select pertinent columns
fundData <- cleaned_cutoff %>% 
  select(Organization,Funding,Tissue) %>% 
  mutate(Funding=round(cleaned_cutoff$Funding, -3))
#create factor levels
lev <- fundData %>% group_by(Organization) %>% summarize(Funding = sum(Funding))
lev <- lev[order(lev$Funding),]
#assign factor levels
fundData$Organization <- factor(fundData$Organization,levels = lev$Organization)
 


#function to find factor levels based on size of bar section 
orderFunction <- function(df,groupBy){
  
}

#filter the bar sections by value
#test <- as_tibble(fundData)
#1. iterate across each group, this one is grouped by organization
#for(i in 1:length(lev$Organization)){
  #2. apply an ordering function inside the loop to get factor levels
#  testLev_perOrg <- test
  
  #3. assign those factors as levels within each group
#  test <- test %>% group_modify(factor)
#}



#plot
bar <- plot_ly(fundData, x = ~Funding, y = ~Organization, type = 'bar',
               name = ~Tissue, color = ~Tissue) %>%
   layout(margin = list(l = 300), title = paste0("Funding per Organization (", titleSnippet, ")"),
                        xaxis = list(title = "Funding ($)"),
                        yaxis = list(title = "Organization Name"), orientation = 'h', barmode='stack')
bar
```

Top 10 most funded organizations
```{r}
fundData <- fundData[order(fundData$Funding, decreasing = T),]
for(i in 1:numPrint){ #print 10 most funded orgs 
  print(as.character(fundData$Organization[i]))
}
```

Bar plot - Amount funding per tissue type
```{r, warning=F, message=F}
tissuesFunded <- cleaned_cutoff %>% select(Funding,Tissue) %>% 
  mutate(Funding = round(Funding,-3)) %>%
  group_by(Tissue) %>% 
  summarize(Funding = sum(Funding))
#factor levels
tissuesFunded$Tissue <- factor(tissuesFunded$Tissue, 
                        levels = tissuesFunded$Tissue[order(tissuesFunded$Funding, decreasing=T)])

bar2 <- plot_ly(tissuesFunded, x = ~Tissue, y = ~Funding, type = 'bar') %>%
   layout(title = paste0("Funding per Tissue Type (", titleSnippet, ")"),
                        xaxis = list(title = "Tissue Type"),
                        yaxis = list(title = "Total Funding ($)"))
bar2
```
Top 5 most funded tissue types
```{r}
tissuesFunded <- tissuesFunded[order(tissuesFunded$Funding, decreasing = T),] #order DF by funding amt
for(i in 1:5){ #print top 5 tissue types 
  print(as.character(tissuesFunded$Tissue[i]))
}
```

Bar plot - Number of grants received per year through 2020
```{r, warning=F, message=F}
years_precutoff <- cleaned %>% select(FY,Tissue) %>% filter(FY!=0)
years_b4tbl <- years_precutoff %>% filter(FY >= cutoffYear)
years <- as.data.frame(table(years_b4tbl), stringsAsFactors = F)
years <- years %>% filter(years$Freq!=0)

hist2 <- plot_ly(x = ~years$FY, y = ~years$Freq, type = 'bar',
                 name = ~years$Tissue, color = ~years$Tissue) %>%
  layout(title=paste0("Number of Awards over Time, per Tissue Type (", titleSnippet, ")"), yaxis = list(tickprefix=" ", title = 'Number of Awards'), xaxis = list(title = 'Year'), barmode = 'stack')
hist2
```

```{r,  warning=F, message=F}
years_cut <- cleaned_cutoff %>% select(FY,Funding,Tissue) %>% filter(FY!=0)
#grouped <- years_cut %>% group_by(FY) %>% summarize(Funding = sum(Funding))

bar4 <- plot_ly(x = ~years_cut$FY, y = ~years_cut$Funding, type = 'bar',
                name = ~years_cut$Tissue, color = ~years_cut$Tissue) %>%
  layout(title=paste0("Amount Funding over Time, per Tissue Type (", titleSnippet, ")"), 
         yaxis = list(tickprefix=" ", title = 'Funding ($)'), xaxis = list(title = 'Year', autotick=F), 
         barmode = 'stack')
bar4

```



