---
title: "SBIR Awards Summary"
author: "Logan Kocka"
date: "6/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(plotly)
library(data.table)
```

User inputs
```{r, message=FALSE}
setwd('S:/Bernard/Logan')
fileSource <- "NSF, all years" #used in title of plots
numPrint <- 10 #prints this many of the highest below plots
```

Load in the data and sort by Organization name.
```{r, message=FALSE}
data <- read.csv('NSF 2020-06-10 armi_df_whole.csv', stringsAsFactors = F) #choose file
data <- data[order(data$Organization),] #order according to Organization (A-Z)
```

Clean the data: remove commas, "Inc.", "Inc", "Llc", and trailing white spaces.
```{r, message=FALSE}
cleaned <- data
cleaned$Organization <- gsub(",", "", cleaned$Organization) #removes commas
cleaned$Organization <- gsub("Incorporated", "", cleaned$Organization) #removes Incorp
cleaned$Organization <- gsub("Inc.", "", cleaned$Organization) #removes Inc.
cleaned$Organization <- gsub("Inc", "", cleaned$Organization) #removes Inc
cleaned$Organization <- gsub("Llc", "", cleaned$Organization) #removes Llc
cleaned$Organization <- trimws(cleaned$Organization) #removes trailing spaces
print(paste0("Number of unique orgs before: ",nrow(data)))
print(paste0("Number of unique orgs after: ",nrow(cleaned)))
print("The above values should be equal.")
```

Filter and organize data for plotting.
```{r, message=FALSE, warning=F}
filtered <- cleaned %>% select(Organization, Tissue) #select rows you want to keep
counts <- as.data.frame(table(filtered), stringsAsFactors = F) 
#consolidate by tissue type and count frequency
counts <- counts[!grepl("0", counts$Freq),] #removes frequencies of zero

#order bars in plot
#remove tissue type column  #sum freq's per org
countsOrder <- counts %>% select(!Tissue) %>% group_by(Organization) %>% summarize(Freq = sum(Freq))
#order by Freq to establish customer factors
countsOrder <- countsOrder[order(countsOrder$Freq, decreasing=T),]
counts$Organization <- factor(counts$Organization, level=countsOrder$Organization)
```

Stacked histogram - Total number of awards per organization
```{r, warning=FALSE}
#create stacked histogram plot showing breakdown of awards per organization and their tissue types
hist <- plot_ly(x = ~counts$Organization, y = ~counts$Freq, type = 'bar', 
                name = ~counts$Tissue, color = ~counts$Tissue) %>%
  layout(title=paste0("Tissue Type / Organization Breakdown (", fileSource, ")"), yaxis = list(tickprefix=" ", title = 'Number of Awards'), xaxis = list(title = 'Organization Name'), barmode = 'stack')
hist
```
Top 10 organizations by most total awards
```{r}
for(i in 1:numPrint){
  print(as.character(countsOrder$Organization[i]))
}
```

Pie plot - Total number of awards per tissue type
```{r}
#create pie plot showing breakdown of total number of awards per tissue type
pieData <- as.data.frame(table(filtered[,-1]), stringsAsFactors = F)
pie <- plot_ly(pieData, labels = ~Var1, values = ~Freq, type = 'pie') %>% 
  layout(title = paste0("Tissue Distribution (", fileSource, ")"),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pie
```
Top 5 tissue types by number of awards
```{r}
pieData <- pieData[order(pieData$Freq, decreasing = T),] #order DF by frequency
for(i in 1:5){ #print top 5 tissue types 
  print(pieData$Var1[i])
}
```

Bar plot - Amount of funding per organization
```{r, warning=F}
#select pertinent columns
fundData <- cleaned %>% select(Organization,Funding)
fundSums <- fundData %>% group_by(Organization) %>% summarize(
  Funding = sum(Funding) #sum funding across all awards by organization
) 
numZero <- nrow(fundSums[fundSums$Funding==0,])

fundSums$Funding <- round(fundSums$Funding, -3)
fundSums$Organization <- factor(fundSums$Organization, 
                                levels = fundSums$Organization[order(fundSums$Funding, decreasing = TRUE)])
bar <- plot_ly(fundSums, x = ~Organization, y = ~Funding, type = 'bar') %>%
   layout(title = paste0("Funding per Organization (", fileSource, ")"),
                        xaxis = list(title = "Organization Name"),
                        yaxis = list(title = "Funding ($)"))
bar
```

Top 10 most funded organizations
```{r}
fundSums <- fundSums[order(fundSums$Funding, decreasing = T),]
for(i in 1:numPrint){ #print 10 most funded orgs 
  print(as.character(fundSums$Organization[i]))
}

print(paste0("The number of awards showing zero funding was ",numZero))
```

Bar plot - Amount funding per tissue type
```{r, warning=F, message=F}
tissuesFunded <- cleaned %>% select(Funding,Tissue)
tissuesFunded$Funding <- round(tissuesFunded$Funding, -3)
tissuesFunded <- tissuesFunded %>% group_by(Tissue) %>% summarize(
  Funding = sum(Funding) #sum funding across all awards by organization
)
tissuesFunded$Tissue <- factor(tissuesFunded$Tissue, 
                        levels = tissuesFunded$Tissue[order(tissuesFunded$Funding, decreasing = TRUE)])

bar2 <- plot_ly(tissuesFunded, x = ~Tissue, y = ~Funding, type = 'bar') %>%
   layout(title = paste0("Funding per Tissue Type (", fileSource, ")"),
                        xaxis = list(title = "Tissue Type"),
                        yaxis = list(title = "Total Funding ($)"))
bar2
```
Top 5 most funded tissue types
```{r}
tissuesFunded <- tissuesFunded[order(tissuesFunded$Funding, decreasing = T),] #order DF by funding amt
for(i in 1:5){ #print top 5 tissue types 
  print(as.character(tissuesFunded$Tissue[i]))
}
```

Bar plot - Number of grants received per year through 2020
```{r, warning=F, message=F}
years <- cleaned %>% select(FY,Tissue) %>% filter(FY!=0)
years <- as.data.frame(table(years), stringsAsFactors = F)
years <- years[!grepl("0", years$Freq),]

hist2 <- plot_ly(x = ~years$FY, y = ~years$Freq, type = 'bar', 
                 name = ~years$Tissue, color = ~years$Tissue) %>%
  layout(title=paste0("Number of Awards over Time, per Tissue Type (", fileSource, ")"), yaxis = list(tickprefix=" ", title = 'Number of Awards'), xaxis = list(title = 'Year'), barmode = 'stack')
hist2
```

```{r}
#SOMETHING TO BE AWARE OF --
nOmit <- nrow(cleaned) - nrow(years)
print(paste0(nOmit, " rows were omitted because they did not capture a year"))
```

